---
title: "gganimate Package"
author: "Benjamin Prieto"
date: "2/15/2022"
output: 
  html_document: 
    keep_md: yes
---

My Focus: HIV and AIDS, disproportional affected groups/demographics over time

# Setup

## Load Packages
```{r, echo = F}
library("tidyverse")
library("ggplot2")
library("janitor")
library("naniar")
library("RColorBrewer")
library("ggthemes")
library("gganimate")
```

## Load Data  

### NYC HIV/AIDS Data

[Data Source: NYC HIV/AIDS Data (2011-2015)](https://data.world/city-of-ny/fju2-rdad)
```{r}
data_raw <- read_csv("data/dohmh-hiv-aids-annual-report-1.csv") 
```

### Data Cleanup and Clean Data Overview

```{r}
data_clean <- data_raw %>% 
  clean_names() %>% 
  na_if("99999") %>% 
  filter(gender != "All",
         borough != "All",
         uhf != "All") %>% 
  select(-uhf, -percent_linked_to_care_within_3_months) #remove variables not of interest
```

*Note: `age` is compiled based on `race` and vice versa, so no analyses can be done looking at the intersection of age and race. Careful with these two variables during analyses.

# Analyses

## Age and HIV

HIV Diagnosis vs. Age Over Time: How does HIV diagnoses change over time across all age groups?

```{r}
data_clean$year <- as.integer(data_clean$year)
```

```{r}
animate(
  data_clean %>% 
    filter(age != "All") %>% 
    select(age, hiv_diagnoses, year) %>% 
    group_by(age, year) %>% 
    summarize(sum_hiv_diagnoses = sum(hiv_diagnoses)) %>% 
    ggplot(aes(x = age, y = sum_hiv_diagnoses, fill = age))+
    geom_col()+
    labs(title = "Year: {frame_time}",
         x = "Age Group",
         y = "HIV Diagnoses")+
    theme_stata()+
    theme(legend.position = "none")+
    transition_time(year)+
    ease_aes("linear"),
  render = gifski_renderer()
)
anim_save(file = "hiv_diagnoses_col")
```

```{r}
animate( 
 data_clean %>% 
    filter(age != "All") %>% 
    group_by(age, year) %>% 
    summarize(sum_hiv_diagnoses = sum(hiv_diagnoses)) %>% 
    ggplot(aes(x = year, y = sum_hiv_diagnoses, color = age))+
    geom_line()+
    labs(x = "Year",
         y = "HIV Diagnoses",
         title = "HIV Diagnoses Over Time by Age Group (2011-2015)",
         color = NULL)+
    theme_stata()+
   transition_reveal(year),
 renderer = gifski_renderer(loop = F)
)
anim_save(file = "hiv_diagnoses_line")
```


AIDS Diagnoses by Age Over Time: Are AIDS diagnoses disproportional across age groups over time?

```{r}
animate(
  data_clean %>% 
    filter(age != "All") %>% 
    select(age, aids_diagnoses, year) %>% 
    group_by(age, year) %>% 
    summarize(sum_aids_diagnoses = sum(aids_diagnoses)) %>% 
    ggplot(aes(x = age, y = sum_aids_diagnoses, fill = age))+
    geom_col()+
    labs(title = "Year: {frame_time}",
         x = "Age Group",
         y = "AIDS Diagnoses")+
    theme_stata()+
    theme(legend.position = "none")+
    transition_time(year)+
    ease_aes("linear"),
  render = gifski_renderer()
)
anim_save(file = "aids_diagnoses_col")
```

```{r}
animate( 
 data_clean %>% 
    filter(age != "All") %>% 
    group_by(age, year) %>% 
    summarize(sum_aids_diagnoses = sum(aids_diagnoses)) %>% 
    ggplot(aes(x = year, y = sum_aids_diagnoses, color = age))+
    geom_line()+
    labs(x = "Year",
         y = "AIDS Diagnoses",
         title = "AIDS Diagnoses Over Time by Age Group (2011-2015)",
         color = NULL)+
    theme_stata()+
   transition_reveal(year),
 renderer = gifski_renderer(loop = F)
)
anim_save(file = "aids_diagnoses_line")
```


HIV-related Death Rate by Age Group over time: Is HIV-related death more or less common across different age groups over time?

```{r}
animate(
  data_clean %>% 
    filter(age != "All", year != 2015) %>% 
    select(age, hiv_related_death_rate, year) %>% 
    group_by(age, year) %>% 
    summarize(avg_hiv_related_death_rate = mean(hiv_related_death_rate)) %>% 
    ggplot(aes(x = age, y = avg_hiv_related_death_rate, fill = age))+
    geom_col()+
    labs(title = "Year: {frame_time}",
         x = "Age Group",
         y = "Average HIV-Related Death Rate (%)")+
    theme_stata()+
    theme(legend.position = "none")+
    transition_time(year)+
    ease_aes("linear"),
  render = gifski_renderer()
)
anim_save(file = "hiv_related-death_rate_col")
```

```{r}
animate( 
 data_clean %>% 
    filter(age != "All", year != 2015) %>% 
    group_by(age, year) %>% 
    summarize(avg_hiv_related_death_rate = mean(hiv_related_death_rate)) %>% 
    ggplot(aes(x = year, y = avg_hiv_related_death_rate, color = age))+
    geom_line()+
    labs(x = "Year",
         y = "Average HIV-Related Death Rate (%)",
         title = "HIV-Related Death Rate Over Time by Age Group (2011-2014)",
         color = NULL)+
    theme_stata()+
   transition_reveal(year),
 renderer = gifski_renderer(loop = F)
)
anim_save(file = "hiv_related_death_rate_line")
```

Comparison with NYC Demographics
```{r}
demo_age_all <- read_csv("data/demo_age_all.csv")
```

```{r}
demo_age_all$age_group <- #order levels by age group instead of alphabetical
  factor(demo_age_all$age_group, 
         levels = c("under 5 years", "5 to 9 years", "10 to 14 years", "15 to 19 years", "20 to 24 years", "25 to 34 years", "35 to 44 years", "45 to 54 years",
                    "55 to 59 years", "60 to 64 years", "65 to 74 years", "75 to 84 years", "85 years and over"))
```

```{r}
demo_age_all$year <- as.integer(demo_age_all$year)
```

```{r}
animate(
  demo_age_all %>% 
    ggplot(aes(x = "", y = percent_pop_for_year, fill = age_group))+
    geom_col(color = "gray50")+
    coord_polar(theta = "y")+
    theme_stata()+
    labs(title = "Year: {frame_time}",
         fill = NULL,
         y = NULL,
         x = "Population")+
    theme(axis.text.x = element_blank(),
          legend.text = element_text(size = 9),
          legend.position = "right")+
    scale_fill_brewer(palette = "Set3")+
    transition_time(year)+
    ease_aes(),
  renderer = gifski_renderer()
)
```